---
title: "Statistical Computing - HW2"
author: "Yaqiong Yao"
date: "9/13/2018"
output: pdf_document 
abstract: "We use this homework to familiar with Rmarkdown and also to learn how computers process numbers, especially the double floating numbers."
header-includes:
- \usepackage{booktabs}
- \usepackage{sectsty} \sectionfont{\centering} \subsectionfont{\centering}
---

# Question 2

Use Monte Carlo method to approximate the distribution function of 

$$ \Phi(t) = \int_{\infty}^t \frac{1}{\sqrt{2\pi}} e^{-y^2/2}dy $$

```{r}
n <- c(1e2, 1e3, 1e4)
t <- c(0.0, 0.67, 0.84, 1.28, 1.65, 2.32, 2.58, 3.09, 3.72)
itr <- 100
set.seed(1)
Phi <- matrix(NA, nrow = length(n), ncol = length(t))
for (ni in seq_along(n)){
  X_ni <- rnorm(n[ni], mean = 0, sd = 1)
  for (ti in seq_along(t)) {
    Phi[ni, ti] <- sum(X_ni <= t[ti])/n[ni]
  }
}
Phi <- rbind(pnorm(t, mean = 0, sd = 1, options(digits = 4)), Phi)
colnames(Phi) <- t
rownames(Phi) <- c("True Value", "100", "1000", "10000")
knitr::kable(Phi)
```

From the table, we know that the monte carlo method works well. The estimated distribution is approximated to the true distribution.

```{r}
# iteration for 100 times
set.seed(12)
Phi <- list()
for (ni in seq_along(n)) {
  Phii <- matrix(NA, nrow = itr, ncol = length(t))
  for (i in 1:itr) {
    X_ni <- rnorm(n[ni], mean = 0, sd = 1)
    for (ti in seq_along(t)) {
      Phii[i, ti] <- sum(X_ni <= t[ti])/n[ni]
    }
  }
  Phi[[ni]] <- Phii
}
# Boxplot
boxplot(Phi[[1]], names = t, xlab = "t", main = "Boxplot when n = 100")
boxplot(Phi[[2]], names = t, xlab = "t", main = "Boxplot when n = 1000")
boxplot(Phi[[3]], names = t, xlab = "t", main = "Boxplot when n = 10000")
```


# Question 3

`.Machine$double.xmax`: The largest normalized floating-point number. It can be expressed by $(-1)^0(1+\sum_{i=1}^{52}2^{-i})\times 2^{2046-1023}$. The first 10 digits of exponent part are 1 and 52 bits of fraction are all 1.
```{r}
u <- 0
for(i in 1:52) u <- u+2^(-i)
(1+u)*2^(1023) == .Machine$double.xmax
```


`.Machine$double.xmin`: The smallest non-zero normalized floating-point number. It can be expressed as $(-1)^0 2^{1-1023}$. Only last bit of exponent is 1 and all 52 bits of fraction are 0.

```{r}
2^{1-1023} == .Machine$double.xmin
```


`.Machine$double.eps`: The smallest positive floating-point number $x$ such that $1+x \ne 1$. It equals to $2^{-52}$.

```{r}
2^(-52) == .Machine$double.eps
```


`.Machine$double.neg.eps`: A small positve floating-point number $x$ such that $1-x \ne 1$. It equals to $2^{-53}$.

```{r}
2^(-53) == .Machine$double.neg.eps
```




